Linearly Ordered Attribute Grammar scheduling using SAT-solving

### 2015-01-06 paper/paper.tex
Fix small typo
Define neighbourhood
Changed my email + some changes + change of LOAG-backtrack compiletime

% Paper info
\title{Linearly Ordered Attribute Grammar scheduling using SAT-solving}
\author{Jeroen Bransen\inst{1}  \and L. Thomas van Binsbergen\inst{2,1} \and Koen Claessen\inst{3} \and Atze Dijkstra\inst{1}}
\institute{Utrecht University, Utrecht, The Netherlands, \email{\{J.Bransen,atze\}@uu.nl}
\and Royal Holloway, University of London, Egham, UK, \email{Thomas.VanBinsbergen.2014@live.rhul.ac.uk}
\and Chalmers University of Technology, Gothenburg, Sweden, \email{koen@chalmers.se}}

@@

\begin{definition}
An undirected graph is \emph{chordal} if and only if it has a 
perfect elimination order. A perfect elimination order is an
ordering $v_1,\ldots,v_n$ of the vertices of $G$ such that in the graph
$G[v_1,\ldots,v_i]$, $\univ{i}{1 \leq i \leq n}$, the vertex $v_i$ is 
\toindex{simplicial}. A vertex $v$ is called simplicial in a graph $G$ if
the neighbourhood of $v$ forms a connected component in $G$.
The graph $G[v_1,\ldots,v_i]$ is the induced subgraph of $G$ containing
only the vertices $v_i,\ldots,v_i$ and the edges between these vertices.
\end{definition}

@@

\subsection{Finding the schedule}
When the constructed Boolean formula is given to the SAT-solver, the result is either that the formula is not satisfiable, meaning that no schedule has been found, of satisfiable, meaning that there is a schedule. It is not hard to see that the formula is satisfiable if and only if there exists a valid schedule for the given attribute grammar definition, and in this paper we give no formal proof of this claim.

@@

\begin{itemize}
\item \emph{Kastens}: the algorithm from \cite{kastens80} which only works for ordered attribute grammars, of which MainAG is not a member. However, with extra manual annotations this algorithm can generate a schedule, which is the approach used previously. However, as writing these annotation is hard an error-prone, this is not an optimal work flow.
\item \emph{Kennedy-Warren}: The algorithm from \cite{kennedywarren76} that we have implemented in the UUAGC before \cite{bransen2012} is an algorithm that schedules all absolutely noncircular attribute grammars, an even larger class than the linearly ordered attribute grammars. However, the scheduling is not completely static so the generated evaluator also does part of the scheduling at runtime. 
\item \emph{LOAG-backtrack}: we have also implemented a backtracking algorithm for linearly ordered attribute grammars \cite{Binsbergen:2015}, based on Kastens' algorithm. This algorithm solves exactly the same scheduling problem as the SAT approach described in this paper and uses exponential time in worst case.
\end{itemize}

\begin{table}
  \begin{center}
    \begin{tabular}{l || c || r}
      Algorithm   &   Manual annotations   &  Compile time MainAG \\
      \hline
      Kastens'                &   Y           &   17s \\
      Kennedy-Warren  &   N           &   33s \\
      LOAG-backtrack  &   N           &   36s \\
      LOAG-SAT           &   N           &   9s \\
    \end{tabular}
  \end{center}
  \caption{Comparison of the four scheduling algorithms}
  \label{tab:algo-comparison}
\end{table}

In \tabref{tab:algo-comparison} we show the compilation times for the MainAG of the UHC for each of the four different algorithms. All times include parsing of the attribute grammar description and code generation, which is what takes most time in the SAT approach. The SAT-solver takes less than a second to find a solution.

>>>>

% Paper info
\title{Linearly Ordered Attribute Grammar scheduling using SAT-solving}
\author{Jeroen Bransen\inst{1}  \and L. Thomas van Binsbergen\inst{2,1} \and Koen Claessen\inst{3} \and Atze Dijkstra\inst{1}}
\institute{Utrecht University, Utrecht, The Netherlands, \email{\{J.Bransen,atze\}@uu.nl}
\and Royal Holloway, University of London, Egham, UK, \email{ltvanbinsbergen@acm.org}
\and Chalmers University of Technology, Gothenburg, Sweden, \email{koen@chalmers.se}}

@@

\begin{definition}
An undirected graph $G$ is \emph{chordal} if and only if it has a 
perfect elimination order. A perfect elimination order is an
ordering $v_1,\ldots,v_n$ of the vertices of $G$ such that in the graph
$G[v_1,\ldots,v_i]$, $\univ{i}{1 \leq i \leq n}$, the vertex $v_i$ is 
\toindex{simplicial}. A vertex $v$ is called simplicial in a graph $G$ if
the neighbourhood of $v$ forms a connected component in $G$.
The neighbourhood of $v$ are all vertices $w$ such that there exists an edge between $v$ and $w$.
The graph $G[v_1,\ldots,v_i]$ is the induced subgraph of $G$ containing
only the vertices $v_i,\ldots,v_i$ and the edges between these vertices.
\end{definition}

@@

\subsection{Finding the schedule}
When the constructed Boolean formula is given to the SAT-solver, the result is either that the formula is not satisfiable, meaning that no schedule can been found, or satisfiable, meaning that there is a schedule. It is not hard to see that the formula is satisfiable if and only if there exists a valid schedule for the given attribute grammar definition, and in this paper we give no formal proof of this claim.

@@

\begin{itemize}
\item \emph{Kastens}: the algorithm from \cite{kastens80} which only works for ordered attribute grammars, of which MainAG is not a member. However, with extra manual annotations this algorithm can generate a schedule, which is the approach used previously. However, as writing these annotation is hard and error-prone, this is not an optimal work flow.
\item \emph{Kennedy-Warren}: The algorithm from \cite{kennedywarren76} that we have implemented in the UUAGC before \cite{bransen2012} is an algorithm that schedules all absolutely noncircular attribute grammars, an even larger class than the linearly ordered attribute grammars. However, the scheduling is not completely static so the generated evaluator also does part of the scheduling at runtime. 
\item \emph{LOAG-backtrack}: we have also implemented a backtracking algorithm for linearly ordered attribute grammars \cite{Binsbergen:2015}, based on Kastens' algorithm. This algorithm solves exactly the same scheduling problem as the SAT approach described in this paper and uses exponential time in worst case. In case of the \emph{MainAG} no backtracking is required to find a schedule.
\end{itemize}

\begin{table}
  \begin{center}
    \begin{tabular}{l || c || r}
      Algorithm   &   Manual annotations   &  Compile time MainAG \\
      \hline
      Kastens'                &   Y           &   17s \\
      Kennedy-Warren  &   N           &   33s \\
      LOAG-backtrack  &   N           &   13s \\
      LOAG-SAT           &   N           &   9s \\
    \end{tabular}
  \end{center}
  \caption{Comparison of the four scheduling algorithms}
  \label{tab:algo-comparison}
\end{table}

In \tabref{tab:algo-comparison} we show the compilation times for the MainAG of the UHC for each of the four different algorithms. All times include parsing of the attribute grammar description, code generation and adding chords, which is what takes most time in the SAT approach. The SAT-solver takes less than a second to find a solution.
